---
title: "AFRE-891 project"
author: "Michael"
date: "2025-04-15"
output: html_document
---

1. Load packages and cluster size data set

```{r}
pacman::p_load(haven, sf, tidyverse, tmap, dplyr, maptiles)

data <- read_dta("/Users/michaelhall/Desktop/school/TZ data/TZ_cleaned_data/N_cluster_size.dta")
```


2. Convert data to data sf

```{r}
# It turns out the gps coordinates were not recorded for some observations!
na_long  <- sum( is.na(data$gps_longitude) )
na_lat   <- sum( is.na(data$gps_latitude) )


# Filter out NAs
data_clean <- data %>%
  filter(
    !is.na(gps_longitude),
    !is.na(gps_latitude)
  )

data_sf <- data_clean |>
  st_as_sf(
    coords = c("gps_longitude", "gps_latitude"),    
    crs    = 4326,               
    remove = FALSE               
  )
```


3. Set CRS for Dar es Salaam, Tanzania 
 
```{r}
data_sf <- st_transform(data_sf, crs = 21037)
```


4. Compute cluster size (a) for 100 meters

```{r}
cluster_list <- st_is_within_distance(data_sf, data_sf, dist = 100)

# Subtract 1 to exclude the vendor themselves from the cluster
data_sf$cluster_100m <- lengths(cluster_list) - 1
```


5. Compute cluster size (b) and (c) for 50 and 20 meters

```{r}
cluster_list <- st_is_within_distance(data_sf, data_sf, dist = 50)
data_sf$cluster_50m <- lengths(cluster_list) - 1

cluster_list <- st_is_within_distance(data_sf, data_sf, dist = 20)
data_sf$cluster_20m <- lengths(cluster_list) - 1
```


Awesome! We now have measures of clusters. Two additional tasks would be:

i. Evaluate the accuracy of vendors' self-reporting
ii. Show the clusters on a map of the city. 

For i. we can compare for each vendor as well as ttest to see if the means are different

```{r}

# Check the names of columns and find the self-reported cluster size of SAME
# food vendors from 2024 (the same year of GPS coordinates)
colnames(data_sf)

# Check the class of the self-reported cluster size to ensure it is numeric
class(data_sf$cluster_size_samefood2024)

# Compare the two cluster measures
mean(data_sf$cluster_size_samefood2024 == data_sf$cluster_100m)

# They are different! Only 8.7 % of vendors reported the same number of adjacent 
# vendors as the 100 meters' calculation.

# An exact number is a high bar. Let's see how many were within 4 vendors.
mean(
  abs(data_sf$cluster_size_samefood2024 - data_sf$cluster_100m) <= 4,
  na.rm = TRUE
)

# Much higher! 58 % of vendors reported a cluster size of vendors within 4 of
# the GPS calculation

# Now ttest to see on the average if the two are different

t.test(
  x      = data_sf$cluster_size_samefood2024,
  y      = data_sf$cluster_100m,
  paired = TRUE,
  alternative = "two.sided"
)

# The t-test suggest the two measures are not significantly different! This is
# great news as it will add robustness to using past measures of cluster size
# obtained via recall.

```


For ii. we begin by creating buffers

```{r}
buffers_100 <- data_sf |>
  st_buffer(dist = 100) |>
  mutate(radius = 100, count = cluster_100m)

buffers_50 <- data_sf |>
  st_buffer(dist = 50) |>
  mutate(radius = 50, count = cluster_50m)

buffers_20 <- data_sf |>
  st_buffer(dist = 20) |>
  mutate(radius = 20, count = cluster_20m)

buffers_all <- bind_rows(buffers_20, buffers_50, buffers_100)
```


Then plot a static map.

```{r}

library(sf)
library(tmap)

# Compute 100 m buffer around each point
buffers_100 <- st_buffer(data_sf, dist = 100)

# Count how many points fall in each buffer
counts <- lengths(st_intersects(buffers_100, data_sf))
buffers_100$cluster_size <- counts

# Plot with bubble size = cluster_size
tmap_mode("plot")
tm_basemap("OpenStreetMap") +
  tm_shape(buffers_100) +
   tm_bubbles(size = "cluster_size",
             fill = "lightblue",
             col = "darkblue",
             lwd = 0.8,
             fill_alpha = 0.3,
             col_alpha = 0.5,
          )    +
  tm_layout(
    frame         = FALSE,
    inner.margins = c(0,0,0,0)
  )


```


Plot an interactive map

```{r}

library(sf)
library(tmap)

# 1. Check data class and CRS
print(class(buffers_100))            
print(st_crs(buffers_100))           

# Transform it to WGS 84 to get it to display properly
buffers_100_ll <- st_transform(buffers_100, crs = 4326)

# 2. Switch to interactive mode
tmap_mode("view")

# 3. Plot again with bubble size = cluster size (and play with styling)
tm_tiles("OpenStreetMap") +
  tm_shape(buffers_100_ll) +
  tm_bubbles(size = "cluster_size",
             fill = "red",
             col = "darkred",
             lwd = 1.2,
             fill_alpha = 0.3,
             col_alpha = 0.5
             )


```


All done! We now have a gps-based measure of 2024 cluster size, we've shown that
street food vendors are able to estimate their cluster within 100m with fairly 
good accuracy, and we can two different maps to help visualize where clusters are
largest within the city. 



